---
title: "Homework_3"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Loading Libraries

```{r}
library(tidyverse)
library(scales)
```


Importing dataset
```{r}
recent_grads <- read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2018/2018-10-16/recent-grads.csv")

recent_grads
```
## Transforming Data
```{r}
recent_grads <- recent_grads %>%
  mutate(Major = str_to_title(Major))

recent_grads
```
## Creating Boxplot
```{r}
recent_grads_plot <- recent_grads %>% 
  ggplot(aes(x = Major_category, y = Median)) + 
           geom_boxplot()

recent_grads_plot
```


## Transposing data so axis are flipped
```{r}
flipped_plot <- recent_grads_plot + coord_flip()

flipped_plot
```


## Adding colors!
```{r}
grads_flipped_colors <- recent_grads %>% 
  ggplot(aes(x = Major_category, y = Median, fill = Major_category)) + 
           geom_boxplot() + coord_flip()
  

grads_flipped_colors
```



## Removing Legend
```{r}
grads_flipped_colors <- recent_grads %>% 
  ggplot(aes(x = Major_category, y = Median, fill = Major_category)) + 
           geom_boxplot() + coord_flip() + theme(legend.position="none")

grads_flipped_colors
```



## Changing x-axis numbers to have dollar signs
```{r}
grads_flipped_colors <- recent_grads %>% 
  ggplot(aes(x = Major_category, y = Median, fill = Major_category)) + 
           geom_boxplot() + coord_flip() + theme(legend.position="none") + scale_y_continuous(labels=scales::dollar_format())

grads_flipped_colors
```



## Changing x-axis scale to start from 0

```{r}
grads_flipped_colors <- grads_flipped_colors + expand_limits(y = 0) # Have to set to y axis because the axis are flipped

grads_flipped_colors
```


## Sort by median ascending

```{r}

grads_flipped_colors <- recent_grads %>% 
  ggplot(aes(x = fct_reorder(Major_category,Median), y = Median, fill = fct_reorder(Major_category, Median))) +
           geom_boxplot() + coord_flip() + theme(legend.position="none") + scale_y_continuous(labels=scales::dollar_format()) + expand_limits(y = 0)

grads_flipped_colors
```


## Changing labels so they are correct
```{r}
grads_flipped_colors <- grads_flipped_colors + labs(y = 'Income (Median)', x = 'Major Category', title = 'Median Income Based on Major Category')

grads_flipped_colors
```


Success!!

# On to question 2 - Bob Ross

Loading library

```{r}
library(janitor)
```


## Importing Data


```{r}
bob_ross <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-08-06/bob-ross.csv")

bob_ross
```


## Clean up column names with Janitor


```{r}
bob_ross <- bob_ross %>% 
  janitor::clean_names()
  
bob_ross
```

## Changing from wide to long format

```{r}
bob_long <- bob_ross %>% 
  pivot_longer(cols = !c('episode','title'),
               names_to = 'element',
               values_to = 'present')

view(bob_long)
```

## Keep only values of n = 1

```{r}
bob_long <- bob_long %>% 
  filter(present == 1)
  
bob_long
```

## Counting and sorting present column

```{r}
bob_long_sorted <- bob_long %>% 
  count(element, sort = TRUE)

bob_long_sorted
```

## Only use the top 20 entries to create bar plot and labeling

```{r}
bob_long_sorted$element <- factor(bob_long_sorted$element, levels = bob_long_sorted$element)

bob_long_sorted %>% 
  head(20) %>% 
  ggplot(aes(x = reorder(element,n), y = n)) +
  geom_bar(stat = 'identity') +
  coord_flip() + 
  labs(x = 'Present in Episode', y = 'In Drawing', title = 'Bob Ross Likes to Draw Trees!!')
  
```



