---
title: "hw4/2"
author: "Matthew Ayala"
date: "11/3/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Loading Libraries 

```{r}
library(caret)
library(skimr)
library(palmerpenguins)
```

# Looking at data

```{r}
penguins
```

# Looking at amount of each species
```{r}
library(tidyverse)
penguins %>% 
  count(species)
```
# Computing mean
```{r}
penguins %>% 
  group_by(species) %>% 
  summarize(across(where(is.numeric), mean, na.rm = TRUE))
```

# Creating validation set

```{r}
validation_index <- createDataPartition(penguins$species, p=0.80, list=FALSE)
# select 20% of the data for validation
validation <- penguins[-validation_index,]
# use the remaining 80% of data to training and testing the models
dataset <- penguins[validation_index,]

```
# Changing columns to double
```{r}
dataset <- dataset %>% 
  mutate(flipper_length_mm = as.numeric(flipper_length_mm)) %>% 
  mutate(body_mass_g = as.numeric(body_mass_g)) %>% 
  select(species,bill_length_mm,bill_depth_mm,flipper_length_mm,body_mass_g)

dataset
```

# Remove NA

```{r}
dataset <- dataset %>% 
  na.omit(dataset)

validation <- validation %>% 
  na.omit(validation)

view(dataset)
```
## Dimensions of of data

```{r}
dim(dataset)
```

#Types of Attributes
```{r}
sapply(dataset, class)
```

# Class Distribution

```{r}
percentage <- prop.table(table(dataset$species)) * 100
cbind(freq=table(dataset$species), percentage=percentage)
```
# Summarize Distribution

```{r}
summary(dataset)
glimpse(dataset)
skim(dataset)
```

# Univariate Plots

```{r}
x <- dataset[,2:5]
y <- dataset[,1]
y
x
```
# Boxplots with the old way
```{r}
par(mfrow=c(1,4))
  for(i in 1:4) {
  boxplot(x[,i], main=names(dataset)[i])
  }
```
# Boxplots with ggplot2
```{r}
things <- c('bill_length_mm','bill_depth_mm','flipper_length_mm','body_mass_g')
long_data <- dataset %>% 
  pivot_longer(cols = things,names_to = 'attribute', values_to = 'values')
ggplot(data = long_data,aes(x=attribute,y=values)) + geom_boxplot() + facet_wrap(.~attribute, scales = "free_y")
```
# Barplot for classes

```{r}
plot(y)

ggplot(data=dataset,aes(x=species)) + geom_bar()
```
# Multivariate Plots

```{r}
y_unl = unlist(y)

featurePlot(x = x,y=as.factor(y_unl), plot = 'ellipse')

```
# Multivariate boxplot

```{r}
featurePlot(x = x,y=as.factor(y_unl), plot = 'box')
```
# Multivariate density plot

```{r}
featurePlot(x = x,y=as.factor(y_unl), plot = 'density',scale = "free")

```

# Test Harness

```{r}
control <- trainControl(method="cv", number=10)
metric <- "Accuracy"
```

# Build models

## lda
```{r}
set.seed(88)
fit.lda <- train(species~., data=dataset, method="lda", metric=metric, trControl=control)
fit.lda
```

## Cart
```{r}
set.seed(88)
fit.cart <- train(species~., data=dataset, method="rpart", metric=metric, trControl=control)
fit.cart
```
## KNN
```{r}
set.seed(88)
fit.knn <- train(species~., data=dataset, method="knn", metric=metric, trControl=control)
fit.knn
```
## SVM

```{r}
set.seed(88)
fit.svm <- train(species~., data=dataset, method="svmRadial", metric=metric, trControl=control)
fit.svm
```

## Random Forest

```{r}
set.seed(88)
fit.rf <- train(species~., data=dataset, method="rf", metric=metric, trControl=control)
fit.rf
```

## Selecting the best model

```{r}
results <- resamples(list(lda=fit.lda, cart=fit.cart, knn=fit.knn, svm=fit.svm, rf=fit.rf))
summary(results)
```

## Comparing accuracy of models

```{r}
dotplot(results)
```
```{r}
print(fit.lda)
```

```{r}
predictions <- predict(fit.lda, validation)
confusionMatrix(predictions, validation$species)
```

