---
title: "Homework 5"
author: "Matthew Ayala"
date: "12/6/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Importing required libraries

```{r}
library(tidyverse)
library(lubridate)
library(here)
library(janitor)
library(apaTables)
library(sjPlot)

```

# Creating list of filesnames

```{r}
filenames <-
  list.files(full.names = TRUE)

filenames
```

# Reading in all the files skipping the first 4 rows.

```{r}
all_files <-
  filenames %>% 
  map(., ~ read_csv(., skip=4))

all_files

```

# Figuring out if tables have the same column names, must read in 4th file skipping only the first 3 column names


```{r}

all_files %>% 
  # Apply the function names to each element of all_files
  map(names)

all_files[[4]] <-
  # Skip the first 3 lines instead of the first 4
  read_csv(filenames[[4]], skip = 3)

all_files
```

# Making sure all the headers are correect now.

```{r}
all_files %>% 
  map(names) %>% 
  map(head)

identical(names(all_files[[1]]), names(all_files[[2]]))
identical(names(all_files[[2]]), names(all_files[[3]]))
identical(names(all_files[[3]]), names(all_files[[4]]))
identical(names(all_files[[4]]), names(all_files[[5]]))

```
# Perfect!

# Now to combine the dataframes.

```{r}
child_counts <-
  all_files %>% 
  bind_rows()

child_counts
```
# Cleaning/ fixing the names with janitor

```{r}

child_counts <- 
  janitor::clean_names(child_counts)

child_counts %>% 
  names()
```
# Showing that we need to clean the state_name column

```{r}
child_counts %>%
  count(state_name) %>%
  head()
```
# Cleaning state_name column and making sure they were all changed
```{r}
child_counts <-
  child_counts %>%
  mutate(state_name = tolower(state_name)) 

child_counts %>%
  count(state_name) %>%
  head()

```
# They were all lowered and now counted as intened!

# Now to change year column data type.
```{r}
child_counts <- child_counts %>%
  mutate(year = ymd(year, truncated = 2))
  

child_counts
```
# Now changing chr columns to numeric

```{r}
child_counts <- child_counts %>% 
  mutate_at(vars(contains('age')),funs(as.numeric))

child_counts
```
# Succesfully processed the data, now on to plotting!

```{r}
# Preselecting columns that we need

child_counts_plotting <- child_counts %>% 
  select(c('year','sea_disability_category','asian_age_3_5'))

child_counts_plotting

```
# Removing NAs and adjusting data to plot
```{r}
child_counts_plotting <- child_counts_plotting %>% 
  filter(sea_disability_category != 'All Disabilities') %>% 
  filter(!is.na(asian_age_3_5)) %>% 
  mutate(sea_disability_category = ifelse(sea_disability_category == 'Developmental delay (valid only for children ages 3-9 when defined by state)', 'Developmental Delay', sea_disability_category))


child_counts_plotting


```
# Creating Plot

```{r}
child_counts_plotting %>% 
  ggplot(aes(x = year, y = asian_age_3_5, fill = sea_disability_category)) +
  geom_bar(stat = 'identity') +
  labs(x ='Year', y = 'Asian Age 3 to 5', title = 'Disability Distribution for Asian')


```
# Now starting with creating the wt4 set.

```{r}
WT4.set <- child_counts %>% 
  filter(
    sea_disability_category == "All Disabilities",
    sea_education_environment %in% c("Total, Age 3-5", "Total, Age 6-21")
  ) %>% 
  select_at(vars(
    year,
    state_name,
    contains("sea"),
    contains("male")
  )) %>% 
  rename(
    state = "state_name",
    age = "sea_education_environment",
    disability = "sea_disability_category",
    f_3_5 = "female_age_3_to_5",
    m_3_5 = "male_age_3_to_5",
    f_6_21 = "female_age_6_to_21",
    m_6_21 = "male_age_6_to_21"
  )

WT4.set
```
# Converting table to long form

```{r}
WT4.set.long <-
  WT4.set %>%
    pivot_longer(cols = f_3_5:m_6_21, 
                 names_to = "gender", 
                 values_to = "total")

WT4.set.long
```



# Making gender columns more intuitive

```{r}
WT4.set.long <- 
  WT4.set.long %>%
  mutate(
    gender = case_when(
      gender == "f_3_5" ~ "f",
      gender == "m_3_5" ~ "m",
      gender == "f_6_21" ~ "f",
      gender == "m_6_21" ~ "m",
      TRUE ~ as.character(gender)
    )
  )

WT4.set.long
```
# Removing NAs

```{r}
WT4.set.long <-
  WT4.set.long %>%
  filter(!is.na(total))

WT4.set.long
```
# Creating high count variable

```{r}
high_count <-
  WT4.set.long %>%
  filter(state %in% c("california", "florida", "new york", "pennsylvania", "texas"))

high_count
```
```{r}
high_count %>%
  filter(gender == "f", age == "Total, Age 6-21") %>%
  ggplot(aes(x = year, y = total, color = state)) +
  geom_freqpoly(stat = "identity", size = 1) +
  labs(title = "Count of Female Students in Special Education Over Time",
       subtitle = "Ages 6-21")
```
# Modeling the data

```{r}
WT4.set.long %>%
  filter(age == "Total, Age 6-21") %>%
  pivot_wider(names_from = gender, 
              values_from = total) %>%
  ggplot(aes(x = f, y = m)) +
  geom_point(size = 3, alpha = .5, color = 'green') +
  geom_smooth() +
  labs(
    title = "Comparison of Female Students to Male Students in Special Education",
    subtitle = "Counts of students in each state, ages 6-21",
    x = "Female students",
    y = "Male students",
    caption = "Data: US Dept of Education"
  )


```
# Creating modeling data set for f/m ratio

```{r}
model_data <- WT4.set.long %>%
  filter(age == "Total, Age 6-21") %>%
  mutate(year = as.factor(year(year))) %>%
  pivot_wider(names_from = gender, 
              values_from = total) %>% 
  # Exclude outliers
  filter(f <= 500000) %>%
  # Compute male student to female student ratio
  mutate(ratio = m / f) %>%
  select(-c(age, disability))

model_data
```
# Plotting the jitter plot

```{r}
ggplot(data = model_data, aes(x = year, y = ratio)) +
  geom_jitter(alpha = .5, color = 'blue') +
  labs(title = "Male to Female Ratio Across Years (Jittered)")
```
# Modeling ratio year

```{r}
ratio_year <- 
  lm(ratio ~ year, data = model_data)

tab_model(ratio_year,
          title = "Table 1")

```
# Final Boxplot

```{r}
model_data %>%
  pivot_longer(cols = c(f, m), 
               names_to = "gender", 
               values_to = "students") %>% 
  ggplot(aes(x = year, y = students, color = gender)) +
  geom_boxplot() +
  scale_y_continuous(labels = scales::comma) +
  labs(
    title = "Median Male and Female Student Counts in Special Education",
    subtitle = "Ages 6-21. Does not include outlying areas and freely associated states",
    x = "",
    y = "",
    caption = "Data: US Dept of Education"
  )
```

